Stage-Splitting a Modal Language

### 2014-07-18 paper/semantics.tex
A few changes to metatheory sections.

%The evaluation judgment is very similar to standard call-by-value evaluation.
%The input to evaluation is a stage-1 expression (usually
%$e$), as well as two administrative contexts ($\Gamma$ and $\Xi$), covered
%later.  Evaluation has two outputs: the {\em future context} (usually $\xi$) and
%the {\em partial value} (usually $v$).  We cover the latter first.  The partial
%value is essentially the result of the first-stage portion of $e$, and must be a
%\bbone-value of the same type as $e$.  As usual, this means that $v$ must be
%composed only of base primitives, tuples, injections, and lambdas (corresponding
%to base types, products, sums, and functions).  Analogously, the value
%corresponding to $\fut$ is a construct called a {\em hatted variable} (written
%$\hat y$), which signifies a reference to some stage-2 computation.  The mapping
%between these hatted variables and stage-two expressions is
%}}}

@@

Because the inputs to evaluation and speculation are sometimes open on variables, we adorn our judgments with two contexts.
The first of these ($\Xi$) keeps track of those hatted variables that may appear in the input.
Since hatted variables can only be introduced to terms via substitution ({\em i.e.} they do not appear in the original program), 
this context is augmented only at stage \bbone\ substitution sites.
The second context ($\Gamma$) keeps track of stage \bbtwo\ variables, which are declared in original program.
% We chose to keep these as two separate contexts, because only $\Xi$ is needed for well-formedness of partial values.

@@

\noindent\TODO terminology for the contexts?

\begin{definition}
Contexts $\Xi$ and $\Gamma$ are well-formed ($\Xi\wf$, $\Gamma\wf$) if they
contain only stage-2 variables.
\end{definition}

\begin{definition}
An environment $\xi$ is well-formed ($\xi\wf$) if either:
\begin{enumerate}
\item $\xi = \cdot$; or
\item $\xi = \xi',x\mapsto e$ where
$\xi'\wf$,
$\Xi,\dom{\xi'};\Gamma\vdash \coltwo{e}{B}$, and
$\Xi,\dom{\xi'};\Gamma\vdash e \res$.
\end{enumerate}
\end{definition}

\begin{theorem}
If $\Xi;\typesone e A$ then $\Xi\wf$, $\Gamma\wf$, and $A\istypeone$.
\end{theorem}

\begin{theorem}
If $\diaonesub$ and $\typesone e A$ then
\begin{enumerate}
\item $\xi\wf$;
\item $\Xi,\dom\xi;\cdot\vdash \colone{v}{A}$; and
\item $\Xi,\dom\xi;\cdot\vdash v\pval$.
\end{enumerate}
\end{theorem}

\begin{theorem}
If $\diatwosub$ and $\typestwo e A$ then
\begin{enumerate}
\item $\Xi;\Gamma\vdash \coltwo{q}{A}$; and
\item $\Xi;\Gamma\vdash q\val$.
\end{enumerate}
\end{theorem}

>>>>

%The evaluation judgment is very similar to standard call-by-value evaluation.
%The input to evaluation is a stage-1 expression (usually
%$e$), as well as two administrative contexts ($\Gamma$), covered
%later.  Evaluation has two outputs: the {\em future context} (usually $\xi$) and
%the {\em partial value} (usually $v$).  We cover the latter first.  The partial
%value is essentially the result of the first-stage portion of $e$, and must be a
%\bbone-value of the same type as $e$.  As usual, this means that $v$ must be
%composed only of base primitives, tuples, injections, and lambdas (corresponding
%to base types, products, sums, and functions).  Analogously, the value
%corresponding to $\fut$ is a construct called a {\em hatted variable} (written
%$\hat y$), which signifies a reference to some stage-2 computation.  The mapping
%between these hatted variables and stage-two expressions is
%}}}

@@

Because the inputs to evaluation and speculation are sometimes open on variables, we adorn our judgments with a context.
Since hatted variables can only be introduced to terms via substitution ({\em i.e.} they do not appear in the original program), 
this context is augmented only at stage \bbone\ substitution sites.
The second context ($\Gamma$) keeps track of stage \bbtwo\ variables, which are declared in original program.

@@

\begin{definition}
Contexts $\Gamma$ is well-formed ($\Gamma\wf$) if it
contains only stage-2 variables.
\end{definition}

\begin{definition}
An environment $\xi$ is well-formed ($\Gamma\vdash\xi\wf$) if either:
\begin{enumerate}
\item $\xi = \cdot$; or
\item $\xi = \xi',x\mapsto e$ where
$\Gamma\vdash\xi'\wf$ and
$\Gamma,\dom{\xi'}\vdash \coltwo{e}{B}$ and
$\Gamma,\dom{\xi'}\vdash e \res$.
\end{enumerate}
\end{definition}

\begin{theorem}
If $\typesone e A$ then $\Gamma\wf$ and $A\istypeone$.
\end{theorem}

\begin{theorem}
If $\diaonesub$ and $\typesone e A$ then
\begin{enumerate}
\item $\Gamma\vdash\xi\wf$;
\item $\Gamma,\dom\xi\vdash \colone{v}{A}$; and
\item $\Gamma,\dom\xi\vdash v\pval$.
\end{enumerate}
\end{theorem}

\begin{theorem}
If $\diatwosub$ and $\typestwo e A$ then
\begin{enumerate}
\item $\Gamma\vdash \coltwo{q}{A}$; and
\item $\Gamma\vdash q\val$.
\end{enumerate}
\end{theorem}
